\nocite{7fc005ba}
\nocite{c7f15065}
The idea of (fiber) bundles is to attach to each point of a (base) space another fixed space - the fiber - yielding a new (total) space over that base space.
But instead of just taking the cartesian product as the total space, one requires the total space to be a cartesian product only locally.
The reason for this more general requirement is that the base space is usually a locally Euclidean space.
The fiber also often has an additional structure, such as that of a vector space.
All in all, these generalized spaces are often appropriate for encoding interesting topological information of a problem - especially in physics - and are therefore well-studied.
The topological information we will later be interested in can be encoded in so-called vector bundles, which we want to discuss now.
\\
A vector bundle is a special type of fiber bundle.
So let us consider fiber bundles first.
A \textbf{fiber bundle} is a tuple $(E,B,p,F)$ consisting of topological spaces $E,B,F$ and a map $p \colon E \to B$ such that there is an open cover $\lbrace U_{\alpha} \colon \alpha \in \mathsf{A} \rbrace$ of $B$ and homeomorphisms $h_{\alpha} \colon p^{-1}(U_{\alpha}) \to U_{\alpha} \times F$ taking $p^{-1}(b_{\alpha})$ to $\lbrace b_{\alpha} \rbrace \times F$ for all $b_{\alpha} \in U_{\alpha}$.
$\lbrace (h_{\alpha},U_{\alpha}) \colon \alpha \in \mathsf{A} \rbrace$ is called a local trivialization, $E$ is the \textbf{total space}, $B$ the \textbf{base space}, and $F$ the \textbf{fiber}.
The fiber bundle $(E,B,p,F)$ is \textbf{trivial} if there is a homeomorphism $p^{-1}(B) \to B \times F$.
Hence, a fiber bundle is a generalization of the cartesian product as a topological space, and cartesian products are a first, though trivial, example of fiber bundles.
\\
Now we can restrict our attention to vector bundles.
They are fiber bundles where the fiber is a specific vector space.
More precisely, a \textbf{(real/complex $n$-dimensional) vector bundle} is a fiber bundle $(E,B,p,\mathbb{K}^{n})$ such that for the local trivialization $\lbrace (h_{\alpha},U_{\alpha}) \rbrace$, the map $h_{\alpha} \vert p^{-1}(b_{\alpha})$ is linear for all $b_{\alpha} \in U_{\alpha}$.
Sometimes one just writes $p \colon E \to B$, or even more simply $E$, for a vector bundle.
It is common as well to write $\epsilon^{n}$ for the $n$-dimensional trivial bundle over a context-related base space.
A non-trivial example is the tangent bundle over the sphere $S^{2}$.
\\
Vector bundles can be reconstructed from their trivial pieces.
The value of this reconstruction comes from making a relation to cohomology apparent, on the one hand.
This is discussed in \cite{catctxphy} and shall not concern us here, although it is also important in our context.
On the other hand, it allows us to construct a vector bundle when only the trivial pieces are given, which is often the case in practice and is thus useful for us.
Now let us look at how the total space of a bundle $(E,B,p,\mathbb{K}^{n})$ can be reconstructed from a local trivialization $\lbrace (h_{\alpha},U_{\alpha}) \rbrace$.
This is done by {\glqq}gluing together{\grqq} the pieces $U_{\alpha} \times \mathbb{K}^{n}$ in the right way.
By gluing, we mean a specific equivalence relation on the disjoint union of trivial spaces $\coprod_{\alpha}U_{\alpha} \times \mathbb{K}^{n}$.
Identifying $(b,x) \in U_{\alpha_{1}} \times \mathbb{K}^{n}$ with $h_{\alpha_{2}}(h_{\alpha_{1}}^{-1}(b,x))$ for $b \in U_{\alpha_{1}} \cap U_{\alpha_{2}}$ yields the equivalence relation that reconstructs $E$ as a quotient space.
Moreover, this identification gives rise to \textbf{gluing functions}, which are the obvious maps $f_{\alpha_{2}\alpha_{1}} \colon U_{\alpha_{1}} \cap U_{\alpha_{2}} \to GL_{\mathbb{K}}(n)$ (in terms of local trivialization $h_{\alpha}$) such that\footnote{This is called the \textit{cocycle condition} and is the link to cohomology.}
\begin{align*}
  f_{\alpha_{3}\alpha_{2}}
  \circ
  f_{\alpha_{2}\alpha_{1}}
  &=
  f_{\alpha_{3}\alpha_{1}}
\end{align*}
on $U_{\alpha_{1}} \cap U_{\alpha_{2}} \cap U_{\alpha_{3}}$.
Here, $GL_{\mathbb{K}}(n)$ denotes the set of invertible linear maps on $\mathbb{K}^{n}$.
These gluing functions already suffice to define vector bundles.
\\
With our understanding of vector bundles as a generalization of a family of vector spaces, it comes as no surprise that we can extend several constructions on vector spaces to vector bundles: direct sums, scalar products, and tensor products.
In the following, we assume $(E_{1},B,p_{1},\mathbb{K}^{n_{1}})$ and $(E_{2},B,p_{2},\mathbb{K}^{n_{2}})$ are vector bundles to define these notions.
\begin{enumerate}
\item[$\bullet$]
  The \textbf{direct sum} of these bundles $E_{1}$ and $E_{2}$ is an $n_{1} + n_{2}$ dimensional bundle with base space $B$ and total space
\begin{align*}
  E_{1} \oplus E_{2}
  &=
  \lbrace
      (v_{1},v_{2})
      \in
      E_{1}
      \times
      E_{2}
    \colon
      p_{1}(v_{1})
      =
      p_{2}(v_{2})
  \rbrace
\end{align*}
That this truly defines such a bundle follows easily from these observations:
\begin{enumerate}
\item[(i)]
the \textbf{restriction} of $p \colon E \to B$ to a subset $\tilde{B}$ of $B$ is a vector bundle $\tilde{p} \colon p^{-1}(\tilde{B}) \to \tilde{B}$, where $\tilde{p} := p \vert p^{-1}(\tilde{B})$
\item[(ii)]
the \textbf{product} of vector bundles $p \colon E \to B$ and $\hat{p} \colon \hat{E} \to \hat{B}$ is a vector bundle $p \times \hat{p} \colon E \times \hat{E} \to B \times \hat{B}$ with fibers $p^{-1}(b) \times \hat{p}^{-1}(\hat{b})$ for $b \in B$ and $\hat{b} \in \hat{B}$.
\end{enumerate}
So restricting the product $E_{1} \times E_{2}$ over the diagonal of $B \times B$ yields $E_{1} \oplus E_{2}$ as a vector bundle.
\item[$\bullet$]
A \textbf{scalar product} is a map $(\cdot\vert\cdot) \colon E_{1} \times E_{1} \to \mathbb{K}$ that is a scalar product (on vector spaces) in each fiber.
One can guarantee the existence of a scalar product if $B$ is compact Hausdorff.
\item[$\bullet$]
For tensor products: keeping the notion of gluing functions in mind, a local trivialization $\lbrace (h_{\alpha}^{i},U_{\alpha}) \rbrace$ implies gluing functions $g_{\alpha_{2}\alpha_{1}}^{i}$ for $(E_{i},B,p_{i},\mathbb{K}^{n_{i}})$.
Then $g_{\alpha_{2}\alpha_{1}}^{1}(b) \otimes g_{\alpha_{2}\alpha_{1}}^{2}(b)$ defines gluing functions for $\coprod_{\alpha}U_{\alpha} \times (\mathbb{K}^{n_{1}} \otimes \mathbb{K}^{n_{2}})$.
The resulting vector bundle is written $E_{1} \otimes E_{2}$ and called the \textbf{tensor product}.
It is again a bundle over $B$.
\end{enumerate}
Interestingly, because scalar products always exist for bundles with a compact Hausdorff base space, one can prove the following proposition, which is the basis for a K-theoretic examination of vector bundles.
\\
\begin{prp}
\label{prp:vbtriv}
If $B$ is a compact Hausdorff space, then for each vector bundle $(E_{1},B,p_{1},\mathbb{K}^{n_{1}})$, there exists a vector bundle $(\hat{E}_{1},B,\hat{p}_{1},\mathbb{K}^{\hat{n}_{1}})$ such that $E_{1} \oplus \hat{E}_{1}$ is trivial.
\end{prp}
In the last part of this subsection, let us look at vector bundles from a categorical viewpoint.
We want to make vector bundles the objects of a category.
The main question then is how to define morphisms.
Since a bundle consists essentially of two spaces - the base and the total space - a morphism should be a pair of functions.
Moreover, the function between total spaces should preserve the structure of the fiber; that is, it should be linear on the fiber.
Therefore given vector bundles $(E_{1},B_{1},p_{1},\mathbb{K}^{n_{1}}),(E_{2},B_{2},p_{2},\mathbb{K}^{n_{2}})$ one defines a \textbf{(vector) bundle map (from $(E_{1},B_{1},p_{1},\mathbb{K}^{n_{1}})$ to $(E_{2},B_{2},p_{2},\mathbb{K}^{n_{2}})$)} as a pair of maps $\varphi \colon B_{1} \to B_{2}$ and $\phi \colon E_{1} \to E_{2}$ such that $\phi$ is linear in each fiber $p^{-1}(b_{1})$ for $b_{1} \in B_{1}$ making the diagram
\begin{align*}
\begin{CD}
  B_{1}
  @>\varphi>>
  B_{2}
  \\
  @Ap_{1}AA
  @AAp_{2}A
  \\
  E_{1}
  @>\phi>>
  E_{2}
\end{CD}
\end{align*}
commute.
Then the real and complex vector bundles are the objects of categories $\mathbb{K}\mathbf{-VB}$, with bundle maps as morphisms.
\\
An interesting property of $\mathbb{K}\mathbf{-VB}$ is that for $(E_{1},B_{1},p_{1},\mathbb{K}^{n})$ and each map $f \colon B_{2} \to B_{1}$, one can obtain a (nearly unique)\footnote{The keyword is universality w.r.t. category theory.} bundle over $B_{2}$ by {\glqq}pulling back{\grqq} $p_{1}$ along $f$ and extending $f$ to a bundle map.
A representation is
\begin{align*}
  p_{2}
  \colon
  E_{2}
  &:=
  \lbrace
      (b_{2},e_{1})
      \in
      B_{2}
      \times
      E_{1}
    \colon
      f(b_{2})
      =
      p_{1}(e_{1})
  \rbrace
  \to
  B_{2}
  ,\qquad
  (b_{2},e_{1})
  \mapsto
  b_{2}
\end{align*}
making $(f,\phi_{f})$ into a bundle map if $\phi_{f} \colon E_{2} \to E_{1}$ denotes the map defined by $\phi_{f}(b_{2},e_{1}) := e_{1}$.
In particular, note that $\phi_{f}$ is an isomorphism from $p_{2}^{-1}(b_{2})$ to $p_{1}^{-1}(f(b_{2}))$ for all $b_{2} \in B_{2}$.
As indicated above, we say that $E_{2}$ is the \textbf{pullback} of $E_{1}$ (along $f$).
\\
With the help of the pullback, we can now define a functor $V_{\mathbb{K}}$ from $\mathbf{Top-cH}$ to $\mathbf{CMon}$, finally setting the stage for topological K-theory.
Let us denote the set\footnote{This is not obvious (compare \cite{7fc005ba}).} of $n$-dimensional vector bundles over a compact Hausdorff base space $B$ by $V_{\mathbb{K}}^{n}(B)$.
This is essentially what $V_{\mathbb{K}}$ will do on objects.
For morphisms, note that from a map $f \colon B_{2} \to B_{1}$, we obtain functions $V_{\mathbb{K}}^{n}(f) \colon V_{\mathbb{K}}^{n}(B_{1}) \to V_{\mathbb{K}}^{n}(B_{2})$ by pulling back along $f$, which satisfy the following properties:
\begin{enumerate}
\item[(F1)]
If $f$ is the identity, the pullback of a bundle along $f$ is the same bundle again, as given by the identity $V_{\mathbb{K}}^{n}(f)$.
\item[(F2$^{\prime}$)]
For maps $f_{ij} \colon B_{i} \to B_{j}$
\begin{align*}
  V_{\mathbb{K}}^{n}(f_{21} \circ f_{32})
  &=
  V_{\mathbb{K}}^{n}(f_{32})
  \circ
  V_{\mathbb{K}}^{n}(f_{21})
\end{align*}
holds.
\end{enumerate}
This turns $V_{\mathbb{K}}^{n}$ into a (contravariant) functor from $\mathbf{Top-cH}$ to $\mathbf{Set}$, at least.
But one can further show that
\begin{align*}
  V_{\mathbb{K}}^{n_{1}+n_{2}}(f)(E_{1} \oplus E_{2})
  &=
  V_{\mathbb{K}}^{n_{1}}(f)(E_{1})
  \oplus
  V_{\mathbb{K}}^{n_{2}}(f)(E_{2})
\end{align*}
as vector bundles.
Thus, by setting
\begin{align*}
  V_{\mathbb{K}}(B)
  &:=
  \bigcup_{n \in \mathbb{N}}
  V_{\mathbb{K}}^{n}(B)
  \\
  V_{\mathbb{K}}(f)
  &:=
  V_{\mathbb{K}}^{n}(f)
\end{align*}
for maps $f \colon B_{2} \to B_{1}$ and $n$-dimensional vector bundles $E_{1}$ over $B_{1}$, it is not too hard to see that $V_{\mathbb{K}}$ is a contravariant functor from $\mathbf{Top-cH}$ to $\mathbf{CMon}$ with respect to the direct sum of vector bundles.
